Modeling and Simulation of Load Balancing Strategies for Computing in High Energy Physics

### 2018-10-02 paper/paper.tex
Add suggestions from review

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{images/resources}
	\caption[]{Estimated CPU resources of ATLAS \cite{community}}
	\label{resources}
\end{figure}

@@

The contribution of this paper is the modeling and simulating of CMS computing jobs at the Tier 1 center GridKa as a proof of concept.
The model and simulation are created using the Palladio simulator. We extended this simulator to be able to simulate load balancing strategies. Furthermore, we created a automated model parameter analysis and model creation, to increase the usability.
This model enables us to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG. Furthermore, it can be used as a foundation for models and simulations of larger parts of the WLCG.

@@

\section{Simulation Process}
\label{sec:process}

\Cref{process} depicts the simulation process starting with matching of the data sets to receiving a valid model of the site and its jobs.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.6\linewidth]{images/process}
	\caption[]{The Simulation Process}
	\label{process}
\end{figure}

@@

After a job is scheduled on a resource container it acquires the specified number of job slots or waits if the node has not sufficient free slots. Afterwards the CPU and I/O demands of the job are processed. To avoid one big I/O operation at the start of a job and to make the processing of I/O more continuously, the processing is modeled using iterations that reflect the number of events. Thus, for each iteration there are I/O and CPU resources needed. Rather than simulating each event, we simulate a fraction of the processed events, because the simulation of each event would harm the scalability of the simulation. 
While processing, the job uses the specified number of cores to process the CPU demand. To simulate not explicitly modeled elements of the jobs such as the overhead created by pilot jobs we added a delay to each job. This delay is calibrated in order to match the measured throughput. After the job finished, it releases its acquired job slots. All of these parameters are specified as a probability function and are extracted from measured performance data as described in \Cref{sec:param}.

@@

To obtain a comprehensive view both of the resource usage of CMS jobs executed at a WLCG grid site as well as the resources available there to process them, combining information from multiple data sources is required:
\begin{itemize}
    \item We acquire Historical job monitoring information from several global CMS monitoring systems: Job reports from JobMonitoring and Workflow and Data Management framework job reports (FWJR documents) are archived on the \emph{analytix} Hadoop cluster.
    The latter are being archived on the cluster as part of the WMArchive project \cite{WMArchiveWorkloadArchive2018}.
    We collect job information using the \emph{CMSSpark} framework by Kuznetsov et al. \cite{cmsspark}.

@@

\section{Validation}
\label{validation}
To validate the model and simulation we simulated CMS jobs at GridKa in April 2018.
The job performance data was randomly separated into two equal sized calibration and test datasets.

@@

\begin{figure}

@@

	\begin{tikzpicture}[scale = 1]

@@

One possible design decision is a more sophisticated load balancing strategy that aims for a better resource utilization of the working nodes by scheduling a balanced share of CPU and IO heavy jobs to it. To simulate this a new scheduling strategy must be implemented.

>>>>

\begin{figure}
	\centering
	\includegraphics[width=0.7\linewidth]{images/resources}
	\caption[]{Estimated CPU resources of ATLAS \cite{community}}
	\label{resources}
\end{figure}

@@

The contribution of this paper is the modeling and simulating of CMS computing jobs at the Tier 1 center GridKa as a proof of concept.
The model and simulation are created using the Palladio simulator, see \Cref{sec:palladio} for details on Palladio. This leads to the research question which modifications are needed so that Palladio can be used to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio. To realise it we created an extension for the simulator that enables us to simulate load balancing strategies and furthermore, we created an automated model parameter analysis and model creation, in order to increase the usability of the simulation.
This model enables us to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG. Furthermore, it can be used as a foundation for models and simulations of larger parts of the WLCG.

@@

\section{Simulation Process}
\label{sec:process}
To receive a valid model of a site and its jobs, we created the simulation process depicted in \Cref{process}. This process consists of five phases: Matching of data sets, grouping the data for job types, analysing the model parameters, generating a blueprint model, and lastly injecting the parameters into this model. Note, only a part of the data is used to calibrate the model, the other part is used for its validation.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.55\linewidth]{images/process}
	\caption[]{The Simulation Process}
	\label{process}
\end{figure}

@@

After a job is scheduled on a resource container it acquires the specified number of job slots or waits if the node has not sufficient free slots. Afterwards, the CPU and I/O demands of the job are processed. To avoid one big I/O operation at the start of a job and to make the processing of I/O more continuously, the processing is modeled using iterations that reflect the number of events. Thus, for each iteration there are I/O and CPU resources needed. Rather than simulating each event, we simulate a fraction of the processed events, because the simulation of each event would harm the scalability of the simulation. 
While processing, the job uses the specified number of cores to process the CPU demand. To simulate not explicitly modeled elements of the jobs such as the overhead created by pilot jobs we added a delay to each job. This delay is calibrated in order to match the measured throughput of the test data. After the job finished, it releases its acquired job slots. All of these parameters are specified as a probability function and are extracted from measured performance data as described in \Cref{sec:param}.

@@

To obtain a comprehensive view both of the resource usage of CMS jobs executed at a WLCG grid site as well as the resources available there to process them, combining information from multiple data sources is required:
\begin{itemize}
    \item We acquire historical job monitoring information from several global CMS monitoring systems: Job reports from JobMonitoring and Workflow and Data Management framework job reports (FWJR documents) are archived on the \emph{analytix} Hadoop cluster.
    The latter are being archived on the cluster as part of the WMArchive project \cite{WMArchiveWorkloadArchive2018}.
    We collect job information using the \emph{CMSSpark} framework by Kuznetsov et al. \cite{cmsspark}.

@@

\section{Validation}
\label{validation}
To validate the model and simulation we simulated CMS jobs at GridKa in April 2018.
The job performance data was randomly separated into two equal sized calibration and test datasets. For the validation we considered the following metrics: Overall throughput of jobs, shares of the job types and the CPU efficiency of the site. These metrics describe how well the simulation matches the measured performance data.

@@

\begin{figure}[h]

@@

	\begin{tikzpicture}[scale = 0.8]

@@

One possible design decision is a more sophisticated load balancing strategy that aims for a better resource utilization of the working nodes by scheduling a balanced share of CPU and I/O heavy jobs to it. To simulate this a new scheduling strategy must be implemented.
